---
title: "Introduction to germseq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{germseq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(germseq)
```

## Introduction
`germseq` is an R package which was created to facilitate a convenient way in order to perform multiple methods of microbial differential abundance analysis (DAA). In particular, we employ three popular methods of analysis: DESeq2, ANCOM-BC, and ALDEx2. With the resulting output, we are able to compare and contrast the output of these methods and evaluate the robustness results.

### Installation

To install this package, start R and enter:
```{r eval=FALSE}
require("devtools")
devtools::install_github("karenliu5/germseq", build_vignettes = TRUE)
library("germseq")
```


## Typical Workflow

Here, we demonstrate a typical workflow with our package, germseq:

### 1 - Clean data
First, we must prepare the data that is to be analyzed. The analysis function of germseq accepts data in the phyloseq-class format. Thus, we will proceed with this tutorial using the `atlas1006` dataset (Lahti et al. 2015) from the `microbiome` package.

```{r message=FALSE}
library(microbiome)
data(atlas1006)
atlas1006
```

From this phyloseq object, notice that there are three tables: the OTU table, the sample data, and the taxonomy table. The experimental data is housed within the Sample Data module. 

```{r message=FALSE}
head(sample_data(atlas1006))
```

As currently the analysis performed by germseq can only compare between two different conditions, here we choose to compare the microbiome of individuals with `bmi_group = obese` and `bmi_group = lean`. Therefore, we subset the sample data table to only contain these two. We will use the `phyloseq` package's `subset_samples` function in order to reflect these changes across all three tables in the phyloseq object.

```{r message=FALSE}
atlas1006_bmi <- phyloseq::subset_samples(atlas1006, (bmi_group == "lean" 
                                                      | bmi_group == "obese"))
head(sample_data(atlas1006_bmi))
```

Thus, with this cleaned dataset, we will continue to the analysis portion.

## 2 - Running the Analysis
In order to run the analysis, we will use the following commands based on the R version that is currently being employed: `compare_DAA_methods()` for R4.0.5 and `compare_DAA_methods_2()` for R4.2.2. These are identical in input and output, and were created due to original limitations during the package development. The parameters for these functions are the following: `ps` a phyloseq object containing the data to analyze, `group` a string refeerring to the grouping variable, `prevThr` the prevalence filter threshold of choice. For more details regarding how these functions work, consult the help documentation using `?compare_DAA_methods` and `? compare_DAA_methods_2`. 

In our example, we use `atlas1006_bmi` as `ps`, `"bmi_group"` as `group`, and the default value of 0.1 as the `prevThr`:

```{r}
summ <- germseq::compare_DAA_methods(ps = atlas1006_bmi, group = "bmi_group", prevThr = 0.1)
head(summ)
```

This will output a summary table containing whether each method found the abundance of each taxon (`taxon`) to be significantly different after adjusting for multiple hypotheses between the two grouping conditions (`aldex2`, `ancombc`, `deseq2`), as well as the number of methods which found the specific taxon to be significantly different (`score`). Using this summary table, we can finally proceed to the visualization step.


## 3 - Visualization
Here, we provide functions in order that takes the results from the previous analysis to visually compare: a) the proportion of taxon found to be significantly different between methods; and b) the proportion of taxon found to be significant by no methods vs. one method vs. two methods vs. all methods. The former provides a sense of the sensitivity of each method, and the latter provides a sense of the robustness of the analysis. 

Thus, we can proceed with the analysis by performing the following.

For a), the proportion of taxon found to be significantly different between methods:
```{r}
visualize_performances_barchart(summ)
```

For b), the proportion of taxon found to be significant by no methods vs. one method vs. two methods vs. all methods
```{r}
visualize_overlap_piechart(summ)
```

For additional help, please consult `?visualize_performances_barchart` and `?visualize_overlap_piechart`. 
